Moment Distances from Robust Subspace for Botnet Detection

### 2019-05-16 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on results

In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries. We define the operator $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.

@@

\begin{figure}[h!]
     \centering
     \includegraphics[width=8cm]{figures/raw_distplot_capture20110812_Dur.png}
     \caption{Distribution of flow duration for scenario 12.}
     \label{fig:fig01}
\end{figure}

@@

\begin{figure}[h!]
     \centering
     \includegraphics[width=8cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
     \caption{Average flow duration for scenario 12, aggregated by 0.15 seconds.}
     \label{fig:fig02}
\end{figure}

The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example. The normal traffic, required for training the semi-supervised approach, is denoted as $\boldsymbol{X}_i^r$

@@

Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise. We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}_i^r$ for computing $\textbf{L}_i^r$, $\textbf{S}_i^r$ and the robust moments, i.e. the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ to these moments and classify as anomalous the observations with the largest distances.

Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparce matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$. This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}. We adopt RPCA with ALM, which can be formulated as

\begin{equation}\label{eq:eq01}
	\min_{\boldsymbol{L,S}}\left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
\end{equation}
where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraint $\boldsymbol{C = D - A - E}$...

@@

\begin{equation}\label{eq:eq02}
	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
\end{equation}

@@

\begin{equation}\label{eq:eq03}
	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
\end{equation}

The general formula of the $u$-th moment can be expressed as
\begin{equation}\label{eq:eq04}
	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
\end{equation}

Therefore, the skewness is calculated by
\begin{equation}\label{eq:eq05}
	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
\end{equation}

and the kurtosis as
\begin{equation}\label{eq:eq06}
	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
\end{equation}

We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$. We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.

The classical Mahalanobis Distance is defined as		
\begin{equation}\label{eq:eq07}
	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
\end{equation}
where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter. The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}. Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA. We also propose to extend the Equation \ref{eq:eq05} to implement a Skewness-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq08}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
\end{equation}

We finally propose to extend the Equation \ref{eq:eq05} to implement a Kurtosis-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq09}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
\end{equation}

The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection. Therefore, the robust subspace learning from normal data combined to MD can be called md-RPCA when using mean-based MD, sd-RPCA when using skewness-based MD, and kd-RPCA when using kurtosis-based MD.

@@

The Algorithm \ref{alg:alg01} presents how the proposed Moment Distances from Robust Subspace for Botnet Detection can be algorithmically described.
\begin{algorithm}
	\label{alg:alg01}
	\SetAlgoLined
	\KwResult{$\boldsymbol{\bar{t}}$}
	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}_i^t$\;
	$\min_{\boldsymbol{L}^n, \boldsymbol{S}^n}\left\|\boldsymbol{L}^n\right\|_* + \lambda\left\|\boldsymbol{S}^n\right\|_1 + \langle \boldsymbol{Y}^n, \boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n\right\|_F^2$\;
	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	\caption{Moment Distances from Robust Subspace}
\end{algorithm}

The steps between 1 and 6 of the Algorithm \ref{alg:alg01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 7 and 12 aim the anomaly detection on new contamianted observations. Note that the steps 7 and 10 refers to the md-RPCA, while the steps 8 and 11 refers to the sd-RPCA, and the steps 9 and 12 refers to the kd-RPCA. It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.

@@

In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach for all botnet malwares of the CTU-13. Therefore, the data set and scenarios for our evaluation can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, ... , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to Table \ref{tab:tab01}.

@@

	The F-measure is the harmonic mean of precision and recall and is defined by 
	\begin{equation}\label{eq:eq10}
		F-measure = 2 * \frac{PR * RC}{PR + RC},				
	\end{equation}
	where $PR$ is the precision, defined by 
	\begin{equation}\label{eq:eq11}
		PR = \frac{True Positive}{True Positive + False Positive}
	\end{equation}
	and $RC$ is the recall, defined by 
	\begin{equation}\label{eq:eq12}
		RC = \frac{True Positive}{True Positive + False Negative}.
	\end{equation}

>>>>

In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries. We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.

@@

\begin{figure}[h!]
     \centering
     \includegraphics[width=10cm]{figures/raw_distplot_capture20110812_Dur.png}
     \caption{Distribution of flow duration for scenario 12.}
     \label{fig:fig01}
\end{figure}

@@

\begin{figure}[h!]
     \centering
     \includegraphics[width=10cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
     \caption{Average flow duration for scenario 12, aggregated by 0.15 seconds.}
     \label{fig:fig02}
\end{figure}

The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example. The normal traffic, required for training the semi-supervised approach, is denoted as $\boldsymbol{X}^r$

@@

Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise. We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}^r$ for computing $\textbf{L}^r$, $\textbf{S}^r$ and the robust moments, i.e. the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ to these moments and classify as anomalous the observations with the largest distances.

Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparse matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$. This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.

According to Wright et al. \cite{wright2009robust}, under rather broad conditions, as long as the error matrix S is sufficiently sparse, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
\begin{equation}\label{eq:eq01}
	(\boldsymbol{\hat{L}}, \boldsymbol{\hat{S}})\leftarrow \min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}
\end{equation}
\begin{center} subject to: $\boldsymbol{X} = \boldsymbol{L} + \boldsymbol{S}$ \end{center}
\begin{equation}\label{eq:eq02}
    \left \| \boldsymbol{L} \right \|_{*} = \sum_{i} \sigma_{i}(\boldsymbol{L})
\end{equation}
\begin{equation}\label{eq:eq03}
    \left \| \boldsymbol{S} \right \|_{1} = \sum_{ij} \left | \boldsymbol{S}_{ij} \right |
\end{equation}
where $\lambda$ is a positive weighting parameter, which determines the sparsity of $\boldsymbol{S}$, $\sigma$ denotes the singular values of a matrix.

Before ALM, some methods were proposed to solve that convex optimization problem, such as Iterative Thresholding (IT) and Accelerated Proximal Gradient (APG). However, according to \cite{lin2010augmented}, both approaches have scalability problems and require a large number of iterations to converge. The Augmented Lagrange Multiplier is proven to have a \emph{Q-linear} convergence speed and experimental results show that ALM is five times faster than APG, which in theory is sub-linear \cite{lin2010augmented}. Furthermore, ALM reaches more accurate results with less iterations.

The RPCA with ALM can be formulated:
\begin{equation}\label{eq:eq04}
	l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}) = \left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
\end{equation}
where $\boldsymbol{Y}$ is the multiplier of the linear constraint, $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}. Thus, an iterative scheme can be presented as:
\begin{equation}\label{eq:eq05}
	\left\{
		\begin{matrix} 
			(\boldsymbol{L}_{k+1}, \boldsymbol{S}_{k+1}) \in \operatorname*{argmin}_{\boldsymbol{L,S} \in \mathbb{R}^{m \times n}} \{l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}, \\ 
			\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k}),
		\end{matrix}
	\right.
\end{equation}

We adopt RPCA with ADM, which, generally speaking, is a practical improvement of the classical Augmented Lagrangian method for solving convex programming problem with linear constraints, by fully taking advantage of its high-level separable structure \cite{yuan2009sparse}. ADM minimizes $\boldsymbol{L}$ and $\boldsymbol{S}$ variables serially by solving the following problems to generate the new iterate:

\begin{equation}\label{eq:eq06}
	\left\{\begin{matrix}
	\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}\\ 
	\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}\\ 
	\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})
	\end{matrix}\right.
\end{equation}

@@

\begin{equation}\label{eq:eq07}
	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
\end{equation}

@@

\begin{equation}\label{eq:eq08}
	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
\end{equation}

The general formula of the $u$-th moment can be expressed as
\begin{equation}\label{eq:eq09}
	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
\end{equation}

Therefore, the skewness is calculated by
\begin{equation}\label{eq:eq10}
	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
\end{equation}

and the kurtosis as
\begin{equation}\label{eq:eq11}
	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
\end{equation}

We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq06} for $\textbf{X}^r$. We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.

The classical Mahalanobis Distance is defined as		
\begin{equation}\label{eq:eq12}
	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
\end{equation}
where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter. The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}. Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA. We also propose to extend the Equation \ref{eq:eq12} to implement a Skewness-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq13}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
\end{equation}

We finally propose to extend the Equation \ref{eq:eq12} to implement a Kurtosis-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq14}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
\end{equation}

The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection. Therefore, the robust subspace learning from normal data combined to MD can be called mmd-RPCA when using mean-based MD, smd-RPCA when using skewness-based MD, and kmd-RPCA when using kurtosis-based MD.

@@

The Algorithm \ref{alg:alg01} presents how the proposed Moment Distances from Robust Subspace for Botnet Detection can be algorithmically described.
\begin{algorithm}
	\label{alg:alg01}
	\SetAlgoLined
	\KwResult{$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$}
	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}^t$\;
	\While{not $\min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}$ of $\boldsymbol{X}^n$}{
		$\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}$\;
		$\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}$\;
		$\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})$\;
	}	
	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	\caption{Moment Distances from Robust Subspace}
\end{algorithm}

The steps between 1 and 10 of the Algorithm \ref{alg:alg01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 11 and 16 aim the anomaly detection on new contamianted observations. Note that the steps 11 and 14 refers to the mmd-RPCA, while the steps 12 and 15 refers to the smd-RPCA, and the steps 13 and 16 refers to the kmd-RPCA. It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.

@@

In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach for all botnet malwares of the CTU-13. Therefore, the data set and scenarios for our evaluation can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, ... , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:tab01}.

@@

	The F-measure is the harmonic mean of precision and recall and is defined by 
	\begin{equation}\label{eq:eq15}
		F-measure = 2 * \frac{PR * RC}{PR + RC},				
	\end{equation}
	where $PR$ is the precision, defined by 
	\begin{equation}\label{eq:eq16}
		PR = \frac{True Positive}{True Positive + False Positive}
	\end{equation}
	and $RC$ is the recall, defined by 
	\begin{equation}\label{eq:eq17}
		RC = \frac{True Positive}{True Positive + False Negative}.
	\end{equation}
